\mytitle

### 2016-08-03 pdocs/thesis/article-section-2.tex
velocity based formulation basics added to article using erleben as source

\section{Adding plasticity to physics engine}

@@

Velocity-based formulation is not often used in structural or mechanical engineering while, 
however, it is popular within physics based game developers and film production teams.
 \citet[p.~45]{erleben.thesis} provides reasoning and theoretical details on why 
velocity-based formulation is  popular in constraint-based rigid body simulation. 
Main reason is that collision handling can be done without additional procedures.
Acceleration-based formulations need to stop at collision and switch to an impulse based method.
In structural analysis, a formulation and associated numerical solution procedure are selected 
based on needed features.
For most complex scenarios finite element method is used.
In most cases, static solution with assumption of linear strain-displacement relation small 
displacement solution using displacement based boundary conditions is used.
\citet{bathe-1975} provides description for handling of various nonlinearities.
In large displacement analysis, formulation may be based on updated formulation (Eulerian) or
Lagrangian formulation where initial configuration is used.
Further enhancements are material nonlinearity and dynamic analysis.
Physics engine provides dynamic analysis with large large reference translations and rotations.

@@

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|c| l| l|}
\hline
{\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
$J_1, \Omega_1$ & Jacobian & m\_J1linearAxis, m\_J1angularAxis \\ 
$J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
$v$ & linear velocity & \\ \hline
$\omega$ & angular velocity & \\ \hline
$c$        &  right side vector   & m\_constraintError \\ \hline
$C$  & constraint force mixing & cfm \\  \hline
$\lambda$ & constraint force &  \\ \hline
$l$ & low limit for constraint force & m\_lowerLimit \\ \hline
$h$ & high limit for constraint force & m\_upperLimit \\ \hline
\end {tabular}
\end{center}
\caption {Constraint parameters} \label{tab:constraintParameters} 
\end {table}

>>>>

% !TeX root = article.tex
\section{Adding plasticity to physics engine}

@@

Velocity-based formulation is not often used in structural or mechanical engineering while, 
however, it is popular within physics based game developers and film production teams.
 \citet[p.~45]{erleben.thesis} provides reasoning and theoretical details on why 
velocity-based formulation is  popular in constraint-based rigid body simulation. 
Main reason is that collision handling can be done without additional procedures.
Acceleration-based formulations need to stop at collision and switch to an impulse based method.

Background
for velocity based formulation shown here is based on  \citet[p.~45-50]{erleben.thesis}.
% pdf page 64
In following section, these formulations will be clarified by simple example using \bullet\ implementation.
Assuming that force $\vec{f}_{true}(t)$, is known the impulse $\vec{J}$
in the time interval $\Delta t $ can we written as
\begin{equation} \label{eq:impulseIntegral}
\vec{J} = \int_{0}^{\Delta t} \vec{f}_{true}(t) dt
\end{equation}
Using Newton's second law of motion $F=ma$ one can solve for the velocity, $\vec{v}^{\Delta t}$
\begin{equation} \label{eq:impulseIntegraWithNewton}
\int_{0}^{\Delta t} m \frac{d\vec{v}}{dt}dt= \int_{0}^{\Delta t} \vec{f}_{true}(t)
\end{equation}
\begin{equation} \label{eq:impulse}
m(\vec{v}^{\, \Delta t} - \vec{v}^{\, 0})=\vec{J}
\end{equation}
where superscripts denote time, i.e. ${v}^{\Delta t}=v(\Delta t)$. Next position can be found
by integrating the velocity.
Derivation of equations of motion is not shown here in detail but results 
for simple contact can be summarized as
equation \ref{eq:eomL} for locations and \ref{eq:eomV} for velocities.
Symbols are summarized in table \ref{tab:eom} and figure \ref{fig:eom}.
\begin{equation} \label{eq:eomL} % pdf page 69
\vec{s}^{\, t+\Delta t} = s^{t}+\Delta t S \vec{u}^{\, t+\Delta t}
\end{equation}

\begin{equation} \label{eq:eomV}
\vec{u}^{\, t+\Delta t} = u^{t}+\Delta t M^{-1}(C N \vec{f}^{\ t+\Delta t} + \vec{f}_{ext}) 
\end{equation}

Friction in contacts and joint constraints can be handled in unified way by refactoring
equation \ref{eq:eomV} to get \ref{eq:eomV2},  \citet[p.~66-67]{erleben.thesis}.
Jacobian terms are derived by taking time derivates of kinematic constraints.

\begin{equation} \label{eq:eomV2}
\vec{u}^{t+\Delta t} = u^{t}+\Delta t M^{-1}(
 J_{contact}^T \vec{\lambda}_{contact}
+ J_{joint}^T \vec{\lambda}_{joint}
+ \vec{f}_{ext})
\end{equation}

\begin{figure}[htb!]
\centering
\begin{tikzpicture}
\coordinate (O1) at(2,1);
\coordinate (O2) at(2,4);
\coordinate (C) at(2,2);
\draw (0,0) -- (4,0) -- (4,2) -- (0,2) --(0,0);
\draw (2,2) -- (3.5,5) -- (0.5,5) -- (2,2) ;
\node at (3.5,0.4) {$B_1$};
\filldraw (O1) circle (0.5mm) node[anchor=north] {$\vec{r}_1$};
\node at (2.8,4.5) {$B_2$};
\filldraw (O2) circle (0.5mm) node[anchor=south] {$\vec{r}_2$};
\filldraw (C) circle (0.5mm) node[anchor=north west] {$\vec{p}_1$};
\draw[-{Stealth[length=3mm]}] (O1) -- (C) node[anchor=north east] {$\vec{r}_{11}$};
\draw[-{Stealth[length=3mm]}] (O2) -- (C) node[anchor=south east] {$\vec{r}_{21}$};
\draw[-latex,thick] (C) -- ++(0,1.4) node[anchor=west] {$\vec{n}_{1}$};
\end{tikzpicture}
\caption{Illustration of nomenclature for equations of motion.}
\label{fig:eom}
\end{figure}

% pdf page 33, notation in typical ODEs
\begin {table}[htb!]
\begin{center}
\begin{tabular}{|l| l|}
\hline
{\bf Symbol} & {\bf Description} \\  \hline
$\vec{r}_i$ & position of center of mass for body i  \\ \hline
$\vec{q}_i$ & orientation as a quaternion for body i  \\ \hline
$\vec{p}_i$ & contact or joint point i  \\ \hline
$\vec{r}_{ki}$ & $\vec{p}_k - \vec{r}_i$  \\ \hline
$\vec{s}$ & $\lbrack \vec{r}_1, \vec{q}_1,...,\vec{r}_n, \vec{q}_n \rbrack ^T $\\ \hline
$S$ & generalized transformation matrix  \\ \hline
$\vec{v}_i$ & linear velocity of  center of mass for body i  \\ \hline
$\vec{\omega}_i$ & angular velocity of center of mass for body i  \\ \hline
$\vec{u}$ & $\lbrack \vec{v}_1, \vec{\omega}_1,...,\vec{v}_n, \vec{\omega}_n \rbrack ^T $\\ \hline
$M$ & generalized mass matrix  \\ \hline
$C$ & contact condition matrix  \\ \hline
$N$ & contact normal matrix  \\ \hline
$J_{contact}$ & Jacobian matrix for contacts  \\ \hline
$\lambda_{contact}$ & vector of lagrange multipliers for contacts  \\ \hline
$J_{joint}$ & Jacobian matrix for joints  \\ \hline
$\lambda_{joint}$ & vector of lagrange multipliers for joints  \\ \hline
\end {tabular}
\end{center}
\caption {Nomenclature for equations of motion} \label{tab:eom}
\end {table}

In structural analysis, a formulation and associated numerical solution procedure are selected 
based on needed features.
For most complex scenarios finite element method is used.
In most cases, static solution with assumption of linear strain-displacement relation small 
displacement solution using displacement based boundary conditions is used.
\citet{bathe-1975} provides description for handling of various nonlinearities.
In large displacement analysis, formulation may be based on updated formulation (Eulerian) or
Lagrangian formulation where initial configuration is used.
Further enhancements are material nonlinearity and dynamic analysis.
Physics engine provides dynamic analysis with large large reference translations and rotations.

@@

\begin {table}[htb!]
\begin{center}
\begin{tabular}{|c| l| l|}
\hline
{\bf Parameter} & {\bf Description} & {\bf btConstraintInfo2 pointer}\\  \hline
$J_1, \Omega_1$ & jacobian & m\_J1linearAxis, m\_J1angularAxis \\
$J_2, \Omega_2$ & & m\_J2linearAxis, m\_J2angularAxis \\ \hline
$v$ & linear velocity & \\ \hline
$\omega$ & angular velocity & \\ \hline
$c$        &  right side vector   & m\_constraintError \\ \hline
$C$  & constraint force mixing & cfm \\  \hline
$\lambda$ & constraint force &  \\ \hline
$l$ & low limit for constraint force & m\_lowerLimit \\ \hline
$h$ & high limit for constraint force & m\_upperLimit \\ \hline
\end {tabular}
\end{center}
\caption {Constraint parameters} \label{tab:constraintParameters} 
\end {table}
